---
title: "Predicting Student Debt"
author: "Georgianna James"
date: '2022-03-02'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Required Packages

```{r packages, message=FALSE}
library(rcfss)
library(tidyverse)
library(tidymodels)
library(rsample)
library(knitr)

theme_set(theme_minimal())
set.seed(100)
```



## Load the data
```{r load data}
rcfss::scorecard
```



### Creating traning and testing sets

```{r q1 split data}

# split data into train and test

scorecard_split <- initial_split(scorecard, prop = 3 / 4)
scorecard_train <- training(scorecard_split)
scorecard_test <- testing(scorecard_split)
```


# Linear Regression Model

```{r q2 lm mod}

# build linear regression model

lm_mod <- linear_reg() %>%
  set_engine(engine = "lm") %>%
  set_mode(mode = "regression")


lm_mod
```
```{r q2 lm fit}

# fit the model

lm_fit <- lm_mod %>%
  fit(debt ~ type + admrate + satavg + cost + netcost + avgfacsal + pctpell + comprate + firstgen + locale, data = scorecard) %>%
  predict(new_data = scorecard_test) %>%
  mutate(true_scorecard = scorecard_test$debt)
```

## Linear Regression Metrics Table

```{r q2 rmse visualize, echo=FALSE}

# calculate rmse of model


rmse <- rmse(lm_fit, true_scorecard, .pred, na_rm = TRUE)

kable(rmse, col.names = c("Metric", "Estimator", "Estimate"))
```



Using a linear model to predict student debt, we get a root mean square error of 3062.022. 


# Linear Regression Model with 10-fold Cross Validation



```{r q3 set folds }

# creating folds

scorecard_fold <- vfold_cv(data = scorecard_train, v = 10)

scorecard_fold
```


```{r q3 cross validation }

# refitting model to samples and collecting metrics

scorecard_fold_pred <- lm_mod %>%
  fit_resamples(debt ~ type + admrate + satavg + cost + netcost + avgfacsal + pctpell + comprate + firstgen + locale, resamples = scorecard_fold) %>%
  collect_metrics()
```

## Linear Regression Model with 10-fold Cross Validation Metrics Table 

```{r q3 visualize}
kable(scorecard_fold_pred, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```



Predicting student debt using a 10-fold cross validation, you get an rmse of 3110.07. 


# Decicion Tree Model

```{r q4 tree model}
tree_mod <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("regression")
```

## Decision Tree Metrics Table 

```{r q4 tree prediction}
tree_pred <- tree_mod %>%
  fit_resamples(debt ~ type + admrate + satavg + cost + netcost + avgfacsal + pctpell + comprate + firstgen + locale, resamples = scorecard_fold) %>%
  collect_metrics()



kable(tree_pred, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```



Predicting student debt using a decision tree model, you get an rmse of 3476.96

# Conclusion

In this file, I have predicted student debt using three different machine learning models, reporting my results in root mean standard error, which is a measure of how concentrated the predictions are around the line of best fit. The models resulted in the following results: 

* Linear regression model: ```rmse``` = 3062.022
* Linear regression model with 10-fold cross validation: ```rmse``` = 3110.07
* Decision tree model: ```rmse``` = 3476.96

The basic linear regression model resulted in the lowest ```rmse``` value of 3062.02
